---
title: "Health Expectancy"
author: "Tim Riffe"
date: "11/20/2020"
output: 
  beamer_presentation:
    colortheme: "frigatebird"

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 

\Large
> Average number of years lived in good health

\pause

> Summary index of an underlying distribution

## Motivation

- Quality vs quantity of years?
- Hidden inequality?
- Process?
- Labor force (GALI question)?
- Advocacy?

## Monitoring

\begin{center}
\includegraphics[width=11cm]{Healthy-life-years-at-birth_04_04_20.png}
\end{center}

## ...

\Large
> What about intervention?

## Campell's Law

> “The more any quantitative social indicator is used for social decision-making, the more subject it will be to corruption pressures and the more apt it will be to distort and corrupt the social processes it is intended to monitor”

\footnotesize
Campbell, D. T. (1979). Assessing the impact of planned social change. Evaluation and program planning, 2(1), 67-90.

## Goodhart's Law

> “Any observed statistical regularity will tend to collapse once pressure is placed upon it for control purposes.”

\footnotesize
Goodhart, Charles (1975). "Problems of Monetary Management: The U.K. Experience". Papers in Monetary Economics. 1. Sydney: Reserve Bank of Australia.

## Anything to discuss?

## Measures

- Sullivan
- Multistate

## Sullivan 
```{r}
pix <- c(0.942832933525023, 0.925366777629796, 0.913878052953308, 0.905779576715131, 
         0.899819586082213, 0.895228598060479, 0.891415053920396, 0.887843431788832, 
         0.883977734168754, 0.879243670749217, 0.872988101030959, 0.864422353013589, 
         0.85253805474243, 0.836000848599391, 0.813416659727759, 0.783638907018477, 
         0.745769572698891, 0.699353730294645, 0.64468449536391, 0.583095062238541, 
         0.5170619612345, 0.449955803009552, 0.385423852181739, 0.326631969263549, 
         0.275728924251826, 0.23377510423577, 0.201092734009695, 0.17782145726523, 
         0.164539328404487, 0.163055237515445)

lx <- c(2, 1.98744304104249, 1.97166480704484, 1.95251099982402, 1.92979047491913, 
  1.90327639864472, 1.87269654546685, 1.83771283501313, 1.7978930617532, 
  1.75267423877303, 1.70131782289022, 1.64285829147167, 1.57605111088635, 
  1.49933944580084, 1.41094952851575, 1.30916546330524, 1.19275620234035, 
  1.06156436022103, 0.917173279730504, 0.763469971998791, 0.606832911763207, 
  0.455676937888572, 0.319219834302956, 0.205583276597247, 0.119696424658899, 
  0.0618521812086196, 0.0278148278482655, 0.0106670648655719, 0.00341876365030296, 
  0.000898335418648003)

a <- seq(50,108,by=2)
# survival curve
plot(a, lx/2, type = 'n', las = 2,
     bg="white",
     col.main="white")
axis(1,col = "white",col.ticks = "white",col.axis ="white")
axis(2,col = "white",col.ticks="white",col.axis ="white",las=1)
# color part of it healthy
lines(a, 1- pix, col = "red", lwd = 2)
lines(a, lx/2, col = "white", lwd = 2)
```



- $HLE(x) = \int l(x)\pi(x) \mathrm{d}x$
- List properties \& biases (5 min)

## Multistate

- Process model
- Draw state space diagrams 
- Draw how it works 

## Multistate

```{r}
library(tidyverse)

TR <- read_csv("TRv6.csv") 
TRsub <- TR %>% 
  filter(sex == "f",
         edu == "terciary",
         time == 1996)

```


## Properties

\begin{center}
\includegraphics[width=11cm]{chuttersnap-gts_Eh4g1lk-unsplash.jpg}
\end{center}

## Properties

\begin{center}
\includegraphics[width=11cm]{josh-withers-3mY4U7FIevc-unsplash.jpg}
\end{center}

## Properties

\begin{center}
\includegraphics[width=11cm]{nii-sx7kn7GNf6E-unsplash.jpg}
\end{center}

## Variants

- Discrete
  - Matrix algebra *
  - Lifetable *
  - Simulation *
- Continuous
  - Matrix algebra
  - Differential equations
  
## Considerations

- Initial conditions?
- Mid-interval approximations?
- Strata or state?
- Memoryless?
- Rate fluctuations?
- Health comparability?
  
## Equivalencies

- Cohorts
- Stationary populations
  - reversibility
  
## Further training

- maybe early December Jutta Gampe Multistate course. Otherwise in 2021
- maaaybe Spring Hal Caswell matrix pop course. 

## Special conferences

- IAPHS <https://iaphs.org/>
- REVES <https://reves.site.ined.fr/en/home/about_reves/>

## Local technical resources

- Transition probs: J. Gampe, A. Carollo, D. Schneider, A. Lorenti, C. Dudel
- Markov measures: C. Dudel, me

## Related things that I work on

- Within-inequality in Sullivan setting
- Decomposition of discrete multistate
- Other within-lifecourse duration measures
- Time perspectives on compression and ageing

## Break, then we code!

\begin{center}
\includegraphics[width=9cm]{is_it_worth_the_time.png}
\end{center}
